---
title: "Homework 2 Problems"
author: "Zicheng Wang"
date: "2024-09-25"
output: github_document
---

Initialization of packages.
```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

Import subway E&T data, clean the names and retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. In addition, the entry has been changed from chr to logic.

```{r}
subway_df = 
  read_csv(file = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv", na = c("NA", "", ".")) |>
  janitor::clean_names() |>
  select(line:entry, vending, ada) |>
  mutate(
      entry = case_match(
        entry,
        "YES" ~ TRUE,
        "NO" ~ FALSE
      )
  )
```

Before the cleaning, the raw data had 32 variables, after the cleaning, the dataset had **19** variables and **1868** observations (dimension: 1868*19). The dataset is tidied according to the instrcutions.

```{r}
distinct_station = nrow(distinct(subway_df, line, station_name))
```

There are **`r distinct_station`** distinct stations in the dataset.

```{r}
subway_df = 
  subway_df |>
  distinct(line, station_name, .keep_all = TRUE)
```

Calculate how many station that are ADA compliant, and the proportion of stations that not support vending in entry/exit.

```{r}
ada_support = sum(subway_df$ada == TRUE)
vending_not_support = sum(subway_df$vending == "NO")
proportion_not_support = vending_not_support / distinct_station * 100
```
There are **`r ada_support`** ADA compliant.
There are **`r proportion_not_support`** % proportion of station that not support vending in entry/exit.

Reformat data so that route number and route name are distinct variables. 

```{r}
subway_tidy_df = 
  subway_df |>
  mutate(
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)
  ) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_names"
  ) |>
  filter(!is.na(route_names))

```

Calculate how many distinct stations serve the A train, and in these station how many are ADA compliant.

```{r}
Station_with_A_route = sum(subway_tidy_df$route_names == "A")
A_route = filter(subway_tidy_df, subway_tidy_df$route_names == "A")
ADA_compliant_A = sum(A_route$ada == TRUE)
```

There are **`r Station_with_A_route`** stations serve the A train, and among these stations, **`r ADA_compliant_A`** stations are ADA compliant.


## Problem 2

Import data of Mr trash wheel
```{r}
mr_trash_wheel_df = 
  read_excel("./202409 Trash Wheel Collection Data.xlsx",
             sheet = "Mr. Trash Wheel",
             range = "A2:N586",
             na = c("NA", "", ".")) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    sports_balls = as.numeric(round(sports_balls, digits = 0)),
    trash_wheel = "mr_trash_wheel",
    year = as.double(year)
  )

```

Show the head of Mr trash wheel
```{r}
head(mr_trash_wheel_df, 5)
```

Import data of professor trash wheel
```{r}
prof_trash_wheel_df = 
  read_excel("./202409 Trash Wheel Collection Data.xlsx",
             sheet = "Professor Trash Wheel",
             range = "A2:M108",
             na = c("NA", "", ".")) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    trash_wheel = "prof_trash_wheel"
  )

```

Show the head of Professor trash wheel
```{r}
head(prof_trash_wheel_df, 5)
```

Import data of gwynnda trash wheel
```{r}
gwynnda_trash_wheel_df = 
  read_excel("./202409 Trash Wheel Collection Data.xlsx",
             sheet = "Gwynnda Trash Wheel",
             range = "A2:L157",
             na = c("NA", "", ".")) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    trash_wheel = "gwynnda_trash_wheel"
  )

```

Show the head of Gwynnda trash wheel
```{r}
head(gwynnda_trash_wheel_df, 5)
```

Combine three existing trash wheel data into one data frame
```{r}
trash_wheel_tidy =
  bind_rows(mr_trash_wheel_df, prof_trash_wheel_df, gwynnda_trash_wheel_df) |>
  janitor::clean_names() |>
  relocate(trash_wheel)
  
```

Calculate the total weight of professor trash wheel and total number of cigarette butts collected by Gwynnda trash wheel in June of 2022.
```{r}
total_weight = sum(trash_wheel_tidy$weight_tons[trash_wheel_tidy$trash_wheel == "prof_trash_wheel"],
                   na.rm = TRUE)

cigarette_butt = sum(trash_wheel_tidy$cigarette_butts[
  trash_wheel_tidy$trash_wheel == "gwynnda_trash_wheel" &
  trash_wheel_tidy$month == "June" &
  trash_wheel_tidy$year == 2022],
  na.rm = TRUE)
```
The total weight of professor trash wheel is **`r total_weight`** tons, and total number of cigarette butts collected by Gwynnda trash wheel in June of 2022 is **`r cigarette_butt`** counts.


## Problem 3

Import data of bakers
```{r}
bakers_df =
  read_csv("./gbb_datasets/bakers.csv", na = c("NA", "", ".")) |>
  janitor::clean_names() |>
  separate(
    baker_name, into = c("baker", "last_name"), sep = " "
  )
```

Import data of bakes
```{r}
bakes_df =
  read_csv("./gbb_datasets/bakes.csv", na = c("NA", "", ".")) |>
  janitor::clean_names() |>
  mutate(
    baker = replace(baker, baker == '"Jo"', "Jo")
  )
```

Import data of results
```{r}
results_df =
  read_csv("./gbb_datasets/results.csv", skip = 2, na = c("NA", "", ".")) |>
  janitor::clean_names()
```

Join the data frame into one using `left_join`

```{r}
baker_show_df = left_join(results_df, bakers_df, by = c("baker", "series")) |>
  left_join(bakes_df, by = c("baker", "series", "episode")) |>
  relocate(baker, last_name, baker_age, baker_occupation, hometown, signature_bake, show_stopper) |>
  arrange(baker) |>
  mutate(result = str_to_lower(result)) |>
  filter(!is.na(result))
```

Export dataframe
```{r}
write_csv(baker_show_df, "./gbb_datasets/baker_show.csv")
```

Show winners and star bakers from session 5 to 10.

```{r}
baker_winner =
  baker_show_df |>
  filter(series >= 5 & series <= 10) |>
  filter(result == "winner" | result == "star baker") |>
  relocate(series, episode, result) |>
  arrange(series, episode)
head(baker_winner, 5)
```

Some star bakers win the runner-up, which exemplify their skills and talent (eg Luis Troyano), while some star bakers were out soon after (Chetna Makan). Also, some winners did not won at least one star baker in previous episode (eg David Atherton), which is a twist for TV show.

Import viewership data
```{r}
viewers_df =
  read_csv("./gbb_datasets/viewers.csv", na = c("NA", "", ".")) |>
  janitor::clean_names()
```

Show the first ten rows of the data

```{r}
head(viewers_df, 10)
```

Calculate average viewership in session 1 and 5

```{r}
session_1 = round(mean(viewers_df$series_1, na.rm = TRUE), 2)
session_5 = round(mean(viewers_df$series_5, na.rm = TRUE), 2)
```

The average viewership in session 1 is **`r session_1`**, and in session 5 is **`r session_5`**.

