---
title: "Homework 2 Problems"
author: "Zicheng Wang"
date: "2024-09-25"
output: github_document
---

Initialization of packages.
```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

Import subway E&T data, clean the names and retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. In addition, the entry has been changed from chr to logic.

```{r}
subway_df = 
  read_csv(file = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv", na = c("NA", "", ".")) |>
  janitor::clean_names() |>
  select(line:entry, vending, ada) |>
  mutate(
      entry = case_match(
        entry,
        "YES" ~ TRUE,
        "NO" ~ FALSE
      )
  )
```

Before the cleaning, the raw data had 32 variables, after the cleaning, the dataset had **19** variables and **1868** observations (dimension: 1868*19). The dataset is tidied according to the instrcutions.

```{r}
distinct_station = nrow(distinct(subway_df, line, station_name))
```

There are **`r distinct_station`** distinct stations in the dataset.

```{r}
subway_df = 
  subway_df |>
  distinct(line, station_name, .keep_all = TRUE)
```

Calculate how many station that are ADA compliant, and the proportion of stations that not support vending in entry/exit.

```{r}
ada_support = sum(subway_df$ada == TRUE)
vending_not_support = sum(subway_df$vending == "NO")
proportion_not_support = vending_not_support / distinct_station * 100
```
There are **`r ada_support`** ADA compliant.
There are **`r proportion_not_support`** % proportion of station that not support vending in entry/exit.

Reformat data so that route number and route name are distinct variables. 

```{r}
subway_tidy_df = 
  subway_df |>
  mutate(
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)
  ) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_names"
  ) |>
  filter(!is.na(route_names))

```

Calculate how many distinct stations serve the A train, and in these station how many are ADA compliant.

```{r}
Station_with_A_route = sum(subway_tidy_df$route_names == "A")
A_route = filter(subway_tidy_df, subway_tidy_df$route_names == "A")
ADA_compliant_A = sum(A_route$ada == TRUE)
```

There are **`r Station_with_A_route`** stations serve the A train, and among these stations, **`r ADA_compliant_A`** stations are ADA compliant.


## Problem 2

Import data of trash wheel
```{r}
trash_wheel_df = 
  read_excel("./202309 Trash Wheel Collection Data.xlsx", na = c("NA", "", ".")) |>
  janitor::clean_names()

```



